---
title: "dplyr_library"
author: "Zane Dax"
output:
  pdf_document:
    toc: yes
  html_document:
    toc: yes
    theme: spacelab
    df_print: tibble
---

# The dplyr library 
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


<style type="text/css">
li {color: #196666; font-size:14;}
a {color: #cc0099; font-size:18;}
p {font-family: monaco;font-size: 18; color:#003333;}
h1 {color:#206040;}
h2 {color:#669900;}
h3 {color:#669900;}
div {color: 'red'; font-size: 18;}
</style>


Load the libraries
```{r}
library(tidyverse)
#install.packages(nycflights13)
library(nycflights13)

```


## Topics covered in this tutorial
* filter()
* arrange()
* select()
* mutate()
* groupby()
* summarize()
* left_join()


```{r}

flights = flights
airlines = airlines
```


### Filter 
filter rows of our datasets based on conditions. The R pipe operator " %>% " is the *then* keyword. The "::" is the function from the library.
```{r filter example}
flights %>% 
  dplyr::filter(month == 1 | month == 2, day==1)

```

then we concat the pipe operator
```{r concat pipe}
flights.filtered = flights %>% 
  dplyr::filter(month %in% c(11, 12)) %>% 
  dplyr::filter(dep_time >= 700) %>% 
  dplyr::filter(carrier != "UA")

flights.filtered
```


### Arrange 
The ``arrange()`` function is used for sorting dataframes, the ``desc()`` helper function used to specify descending rather than ascending order. SQL the sort_by.
```{r arrange function}
flights.arranged = flights.filtered %>% 
  arrange( year, month, day, desc(dep_delay))

flights.arranged
```


### select
The select function is used for selecting variables (columns) of the dataset. The use of '-' is to remove column, and ``rename()`` can rename column names.

```{r select function}
flights.selected = flights.arranged %>% 
  dplyr::select(-(hour:time_hour)) %>% # dropping hour from time_hour column
  dplyr::rename("flight_time" = "air_time", "destination" = "dest")

flights.selected
```  


### Mutate
The mutate function can create new variables
```{r mutate}
flights.mutated = flights.selected %>% 
  mutate(gain = dep_delay - arr_delay,
         hours = flight_time/60,
         gain_per_hour = gain/hours)

flights.mutated
```


there are mutate extensions
```{r mutate extensions}
flights.2a = flights.mutated %>% 
  mutate_at(.vars = c("year","month","day"), .funs = as.character)

flights.2b = flights.mutated %>% 
  mutate(across(.cols = c("year","month","day"), .fns = as.character ))
flights.2b

```


### Groupby, summarize
The groupby() function creates a "grouped" dataset, it is used with summarize() function.
```{r Groupby}

mean.delays = flights.mutated %>% 
  group_by(carrier) %>% 
  dplyr:: summarize(mean_delay = mean(dep_delay, na.rm= TRUE)) %>% 
  arrange( desc(mean_delay))

mean.delays
```

the count function is simply ``n()`` or ``tally()``
```{r tally}

carrier.counts = flights.mutated %>% 
  group_by(carrier) %>% 
  dplyr::summarise(n = n()) %>% 
                     arrange(desc(n))

carrier.counts
```


### Joins
All the joins you could ever want
```{r joins}
airline.names = mean.delays %>% 
  left_join(airlines, by= c("carrier" = "carrier")) %>% 
  dplyr::select(name, carrier, mean_delay)

airline.names
```




















