---
title: "Flexdashboard 1"
output: flexdashboard::flex_dashboard
theme: 
  version: 4
  bg: "#101010"
  fg: "#FDF7F7" 
  primary: "#ED79F9"
  navbar-bg: "#3ADAC6"
orientation: columns
runtime: shiny
---




```{r,setup, include = FALSE, context = "server"}
library(bslib)
library(thematic)

# shiny::bootstrapLib()
# bslib::bs_theme()
# bs_theme()
# thematic::thematic_rmd()
# solar_theme <- bs_theme(
#   , , p,
#   base_font = font_google("Pacifico")
# )

# thematic_shiny()

# bs_theme_preview(bs_theme(bg = "#444444", fg = "#E4E4E4", primary = "#E39777"))

```






```{r,  include=FALSE}
library(tidyverse)
library(DT)
startData <- starwars
GBChoices <- as.list(names(startData))
names(GBChoices) <- paste(names(startData),map(startData,~length(unique(.x))))
updateData <- reactive(
  startData %>% group_by(!!! rlang::syms(input$GB)) %>% summarise_if(is.numeric,sum,na.rm=T))
```


















Page 1
====================

Column {.sidebar} 
------------------------------------------------------------------
```{r}
# inputs go here 

selectInput(inputId = "GB",
            label = "Group By",
            choices = GBChoices)

selectInput(inputId = "Metric",
            label = "Metric",
            choices = names(select_if(startData,is.numeric)))
```

Column 
------------------------------------------------------------------




### chart 1 {data-pading=30}
```{r fig.height=15, fig.width=5}
renderPlot({
 updateData() %>% 
  ggplot(aes(y=!! rlang::sym(input$GB),
             x=!! rlang::sym(input$Metric),
             fill=!! rlang::sym(input$GB))) +
  geom_col()+
    theme(legend.position = "none")
})
```





Page 2
====================


### Chart 2  {data-pading=30}
```{r}
starwars %>% 
  select(name, homeworld) %>% 
  count(homeworld, sort = T) %>% 
  ggplot( aes(x= n, y=  reorder(homeworld, n )) )+
  geom_col(fill='#b9d11d')+
  ggdark::dark_mode()+
  labs(title = "Starwars count of homeworlds by name")
```






Page 3 {data-icon="fa-hashtag"}
====================

### Table  {data-pading=30}
```{r}
renderDT(
  updateData(), rownames = F, extensions = 'Buttons', filter="top", editable=T,
  options = list(
    dom = 'Blfrtip',
    buttons = c('copy', 'csv', 'excel', 'pdf', 'print'),
    lengthMenu = list(c(10,50,100,-1),c(10,50,100,"All"))
  )
)
```





page 4 {data-icon="fa-list"}
=================
```{r , include=FALSE}
library(dygraphs)
library(flexdashboard)
```

### Lung Deaths (All)

```{r}
dygraph(ldeaths)
```

### Lung Deaths (Male)

```{r}
dygraph(mdeaths)
```

### Lung Deaths (Female)

```{r}
dygraph(fdeaths)
```



Page 5
=============
i put words here 

### Cars

i put words here too!
```{r}
knitr::kable(mtcars)
```


### Cars 2

```{r}
renderTable({
  head(mtcars, n = input$rows)
})
```




Page 11 {data-navmenu="Menu A"}
=====================================
You can link to a dashboard page with either of the following syntaxes:

[Page 2]

[Page Two](#page-2)

The second syntax is used when you want a custom name for the link 
(rather than just using the page title).

Page 12 {data-navmenu="Menu A"}
=====================================  


Page 13 {data-navmenu="Menu B"}
=====================================


Page 14 {.storyboard}
=====================================  

### Frame 1 heey

Starwars Eye Colors
```{r}
knitr::kable(starwars$eye_color) 
```

***
comments go here



### Frame 2 over here is the stuff
more data analysis things 
```{r}
knitr::kable(starwars$name) 
```

Details
=========================================

Column
-----------------------------------------







